---
title: "analysis"
author: "Jon Ehrenberg"
date: "2023-10-01"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load packages
```{r message=FALSE, warning=FALSE}
library(raster)
library(habtools)
library(rgdal)
library(plot3D)
library(dplyr)
library(tidyverse)
library(tidyr)
```

Calculate RDH for South dems
```{r}

#get folder path with rasters
path_s <- "C:/Users/Jgke8/ASSE_complexity/data/dems/South/"

#list all files
files_s <- list.files(path_s, pattern = "tif$", full.names = TRUE)

#create empty dataframe to be filled in with RDH values
S_lagoon <- data.frame()

#for loop
for (f in files_s){
  #load one raster
  r <- raster(f)
  #calculate RDH
  habcomp <- rdh(r, lvec = c(0.0625, 0.125, 0.25, 0.5, 1, 2))
  #get file name
  ID <- tools::file_path_sans_ext(f)
  #add file name to habcomp data frame
  habcomp <- habcomp %>% mutate(ID = ID)
  #add values to data frame
  S_lagoon <- rbind(S_lagoon, data.frame(ID=habcomp$ID, R=habcomp$R, D=habcomp$D, H=habcomp$H))
}

S_lagoon
```

Clean up S_lagoon
```{r}
S_lagoon$ID <- gsub("C:/Users/Jgke8/ASSE_complexity/data/dems/South/", "", S_lagoon$ID)
S_lagoon$ID <- gsub("Bommie_", "", S_lagoon$ID)
```


Calculate RDH for North dems
```{r}

#get folder path with rasters
path_n <- "C:/Users/Jgke8/ASSE_complexity/data/dems/North/"

#list all files
files_n <- list.files(path_n, pattern = "tif$", full.names = TRUE)

#create empty dataframe to be filled in with RDH values
N_lagoon <- data.frame()

#for loop
for (f in files_n){
  #load one raster
  r <- raster(f)
  #calculate RDH
  habcomp <- rdh(r, lvec = c(0.0625, 0.125, 0.25, 0.5, 1, 2))
  #get file name
  ID <- tools::file_path_sans_ext(f)
  #add file name to habcomp data frame
  habcomp <- habcomp %>% mutate(ID = ID)
  #add values to data frame
  N_lagoon <- rbind(N_lagoon, data.frame(ID=habcomp$ID, R=habcomp$R, D=habcomp$D, H=habcomp$H))
}

```

Clean up N_lagoon
```{r}
N_lagoon$ID <- gsub("C:/Users/Jgke8/ASSE_complexity/data/dems/North/", "", N_lagoon$ID)
N_lagoon$ID <- gsub("Bommie_", "", N_lagoon$ID)
```

Make full data frame
```{r}
All_data <- rbind(N_lagoon, S_lagoon)

summary(All_data)

```


Bin the bommies into high and low complexity (Not useful in later analyses)
```{r}
#get mean fractal dimension to use as boundary for binning
summary(All_data$D)


All_data <- within(All_data, {
  complexity.D <- NA
  complexity.D[D < 2.765]           <- "L"
  complexity.D[D > 2.765] <- "H"
  
})
```

Label each bommie based on survey area
```{r}
library(quadcleanR)

All_data <- categorize(data = All_data, column = "ID", values = c("North", "South"), name = "Location", binary = FALSE, categories = c("N", "S"), exact = FALSE)
```

Write to a CSV file for data input
```{r}
write.csv(All_data, file = "output/All_data_comp.csv")
```


Run normality on fractal dimension data
```{r}
shapiro.test(All_data$D)
#looks good: W = 0.97847, p-value = 0.3062
```

Run 2-sample t-test
```{r}
t.test(D ~ Location, data = All_data)

#t = 2.0687, df = 62.46, p-value = 0.04272

#The mean D and SE between N and S sites are statistically different
library(plotrix)

mean_D <- All_data %>% select(Location, D) %>% group_by(Location) %>% summarise_if(is.numeric, mean, na.rm = TRUE)


SE_D <- All_data %>% group_by(Location) %>% summarise(se_D=plotrix::std.error(D))

sum_D <- mean_D

sum_D$SE <- SE_D$se_D


#mean_D <- mean_D %>% pivot_longer(!Location, names_to = )

#library(data.table)
#All_data1 <- setDT(All_data)
#All_data1 <- All_data1[, c("mean", "SE") :=list(mean(D), plotrix::std.error(D)), by = Location]


```


Make some plots
```{r}
library(ggplot2)

Dt <- paste("t == 2.07")
Dp <- paste("p == 0.04")

ggplot(All_data, aes(x = Location, y = D, fill = Location)) + geom_boxplot() + ylab("Fractal Dimension (D)")+ theme(panel.border = element_rect(colour = "black", fill = NA, linewidth = 1), panel.background = element_rect(fill = "white")) + annotate("text", x = 2.3, y = 2.93, label = Dt, parse = TRUE) + annotate("text", x = 2.3, y = 2.9, label = Dp, parse = TRUE)


ggplot(sum_D, aes(x = factor(Location), y = D)) +
  geom_bar(position = "dodge", stat = "identity", color = "black", fill = "white") +
  geom_errorbar(aes(ymin = D - SE, ymax = D + SE), width = 0.5) +
  scale_y_continuous(limits = c(2, 3), oob = scales::squish) +
  ylab("Mean fractal dimension (D)") +
  xlab("Study site") +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    panel.background = element_rect(fill = "white")
  ) +
  annotate("text", x = 2.4, y = 2.97, label = Dt, parse = TRUE) +
  annotate("text", x = 2.4, y = 2.9, label = Dp, parse = TRUE) +
  guides(fill = guide_legend(title = "Study site"))



```


Load field observations
```{r}
obs_S <- read.csv("data/Field-obs/10-05-2023_South.csv")
obs_N <- read.csv("data/Field-obs/10-06-2023_North.csv")

```


```{r}
All_field <- rbind(obs_N, obs_S)
```

Stats on Alert and Flight Initiation Distances, pred vs non-pred models (South Site)
```{r}

shapiro.test(obs_S$P_AD.m.)
shapiro.test(obs_S$NP_AD.m.)
shapiro.test(obs_S$P_FID.m.)
shapiro.test(obs_S$NP_FID.m.)


#data transformations not working move to Mann-Whitney

#Alert distance predator vs non-predator model
wilcox.test(obs_S$P_AD.m.,obs_S$NP_AD.m., paired = FALSE)

#W = 178, p-value = 0.0495
#medians are statistically different p-value < alpha (0.05)


#Flight initiation distance pred vs non-pred model
wilcox.test(obs_S$P_FID.m., obs_S$NP_FID.m., paired = FALSE)

#W = 173.5, p-value = 0.0248
#medians are statistically different

```


Stats on Alert and Flight Initiation Distances, pred vs non-pred models (North Site)
```{r}

shapiro.test(obs_N$P_AD.m.)
shapiro.test(log(obs_N$NP_AD.m.))
shapiro.test(obs_N$P_FID.m.)
shapiro.test(log(obs_N$NP_FID.m.))


#data transformations not working move to Mann-Whitney

#Alert distance predator vs non-predator model
wilcox.test(obs_N$P_AD.m.,obs_N$NP_AD.m., paired = FALSE)

#W = 873.5, p-value = 0.002175
#medians are statistically different p-value < alpha (0.05)


#Flight initiation distance pred vs non-pred model
wilcox.test(obs_N$P_FID.m.,obs_N$NP_FID.m., paired = FALSE)

#W = 891.5, p-value = 0.001053
#medians are statistically different

```

Stats on Alert and Flight Initiation Distances, pred vs non-pred models (All)
```{r}
shapiro.test(All_field$P_AD.m.) #normal
shapiro.test(All_field$P_FID.m.) #normal
shapiro.test(All_field$NP_AD.m.) #normal
shapiro.test(All_field$NP_FID.m.) #non-normal

t.test(All_field$P_AD.m., All_field$NP_AD.m., paired = FALSE)
#t = 3.6229, df = 90.797, p-value = 0.0004798 
#Mean AD between P and NP are statistically different

#Mann-Whitney (sample sizes of two groups are not equal)
FID_wil <- wilcox.test(All_field$P_FID.m., All_field$NP_FID.m., paired = FALSE)
#W = 1817.5, p-value = 0.0002076
#Median FID between P and NP are statistically different

```

Plots: boxplot and bar
```{r}
#box for FID
ggplot(All_field) + geom_boxplot(aes(x = "Non-predator", y = NP_FID.m.), fill = "white", color = "black") + geom_boxplot(aes(x = "Predator", y = P_FID.m.), fill = "white", color = "black") + ylab(bquote(Flight~initiation~distance~(m))) + xlab(bquote(Model~type))+ theme(panel.border = element_rect(colour = "black", fill = NA, linewidth = 1), panel.background = element_rect(fill = "white")) + annotate("text", x = 2.3, y = .78, label = "paste(W,\" = 1817.5\")", parse = TRUE) + annotate("text", x = 2.3, y = .73, label = "paste(p,\" = 2.08e-4\")", parse = TRUE) + theme(legend.position = "none")

#bar for AD
#Data prep
mean_AD <- All_field %>% select(NP_AD.m., P_AD.m.) %>% summarise_if(is.numeric, mean, na.rm = TRUE)
names(mean_AD)[colnames(mean_AD) == "NP_AD.m."] <- "Non-predator"
names(mean_AD)[colnames(mean_AD) == "P_AD.m."] <- "Predator"

rotated_df <- pivot_longer(mean_AD, cols = everything(), names_to = "Model_type", values_to = "Mean_dist")

SE_NP_AD <- sd(All_field$NP_AD.m., na.rm = TRUE)/sqrt(sum(!is.na(All_field$NP_AD.m.)))

SE_P_AD <- sd(All_field$P_AD.m., na.rm = TRUE)/sqrt(sum(!is.na(All_field$P_AD.m.)))

SE_AD <- data.frame(SE = c(SE_NP_AD, SE_P_AD))

sum_AD <- rotated_df

sum_AD$SE <- SE_AD$SE

#t = 3.6229, df = 90.797, p-value = 0.0004798 
#Plot
ggplot(sum_AD, aes(x = factor(Model_type), y = Mean_dist)) +
  geom_bar(aes(fill = "white"), position = "dodge", stat = "identity", color = "black") +
  geom_errorbar(
    aes(ymin = Mean_dist - SE, ymax = Mean_dist + SE),
    width = 0.5,
    position = position_dodge(width = 0.75)
  ) +
  scale_fill_manual(values = "white") +  # Set the fill color to white
  scale_y_continuous(limits = c(0, 0.4), oob = scales::squish) +
  ylab("Mean alert distance (m)") +
  xlab("Model type") +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    panel.background = element_rect(fill = "white"),
    legend.position = "none"  # Remove the legend
  ) +
  annotate("text", x = 2.25, y = 0.38, label = "paste(t,\" = 3.62\")", parse = TRUE) +
  annotate("text", x = 2.3, y = 0.35, label = "paste(p,\" = 4.80e-4\")", parse = TRUE)

```


Look at D on predator model response distances
```{r}
#Re-check comparison of D between sites

#Remove missing data from south site obs
obs_S_r <- na.omit(obs_S)

#test normality
shapiro.test(obs_S_r$D) #not normal
shapiro.test(obs_N$D) #not normal

#compare medians
wilcox.test(obs_S_r$D, obs_N$D, paired = FALSE)

#W = 58, p-value = 0.0002624
#Medians are statistically different
```

```{r}
#make a plot

ggplot(All_field, aes(x = Location, y = D, fill = Location)) + geom_boxplot() + ylab(bquote(Fractal~dimension~(D))) + theme(panel.border = element_rect(colour = "black", fill = NA, linewidth = 1), panel.background = element_rect(fill = "white")) + annotate("text", x = 2.3, y = 2.93, label = "paste(W,\" = 58\")", parse = TRUE) + annotate("text", x = 2.38, y = 2.9, label = "paste(p,\" = 2.62e-4\")", parse = TRUE)

```



```{r}
#Add Area and perimeter data

S.A <- read.csv("data/gis/South_PA.csv")
N.A <- read.csv("data/gis/North_PA.csv")

All.A <- rbind(S.A, N.A)

All_field.1 <- merge(All_field, All.A, by=c("ID", "Location"))

#Remove rows with missing data
All_field_D <- na.omit(All_field.1[-c(6,13:17)])

```


```{r}
shapiro.test(All_field_D$D)
#W = 0.98112, p-value = 0.5618
#Fractal dimension data are non-normal

AD_All <- lm(data = All_field_D, log(P_AD.m.) ~ D)
summary(AD_All)
#no relationship between fractal dimension and Alert Distance


FID_All <- lm(log(P_FID.m.) ~ D, data = All_field_D)
summary(FID_All)
#relationship between fractal dimension and Flight Initiation Distance
#Multiple R-squared:  0.1171,	Adjusted R-squared:  0.09978 
#F-statistic: 6.764 on 1 and 51 DF,  p-value: 0.01214

```

Make some plots for regressions
```{r}
#Alert distances
ggplot(All_field_D, aes(D, log(P_AD.m.))) + geom_point(size=3) + geom_smooth(method = "lm", se = FALSE, color = "black")+ ylab("log(Alert distance (m))") + xlab("Fractal dimension") + theme(panel.border = element_rect(colour = "black", fill=NA, linewidth = 1), panel.background = element_rect(fill = 'white')) + annotate("text", x = 2.92, y = -.7, label = "paste(r ^ 2, \" = 0.004\")", parse = TRUE) + annotate("text", x = 2.92, y = -.85, label = "paste(p, \" = 0.273\")", parse = TRUE)


#Flight Initiation distances
lrFID <- paste("r^2 == 0.099")
lpFID <- paste("p == 0.012")

ggplot(All_field_D, aes(D, log(P_FID.m.))) + geom_point(size=3) + geom_smooth(method = "lm", se = FALSE, color = "black")+ ylab(bquote(log(Flight~initiation~distance~(m)))) + xlab(bquote(Fractal~dimension~(D))) + theme(panel.border = element_rect(colour = "black", fill=NA, linewidth = 1), panel.background = element_rect(fill = 'white')) + annotate("text", x = 2.92, y = -.3, label = "paste(r ^ 2, \" = 0.099\")", parse = TRUE) + annotate("text", x = 2.92, y = -.45, label = "paste(p, \" = 0.012\")", parse = TRUE)
```


PCA analyses

Normalize data
```{r}
library(corrr)
library(ggcorrplot)
library(FactoMineR)
library(factoextra)


All.field.num <- All_field_D[,c(3:8, 10, 12:13)]
data.norm <- scale(All.field.num)

head(data.norm)


#Correlation matrix
corr_matrix <- cor(data.norm)
ggcorrplot(corr_matrix)

```


Applying PCA
```{r}
data.pca <- princomp(corr_matrix)
summary(data.pca)


data.pca$loadings[, 1:2]

#Plot eigan values
fviz_eig(data.pca, addlabels = TRUE)

# Graph of the variables
fviz_pca_var(data.pca, col.var = "black")

fviz_cos2(data.pca, choice = "var", axes = 1:2)

fviz_pca_var(data.pca, col.var = "cos2",
            gradient.cols = c("black", "orange", "green"),
            repel = TRUE)


```

Regression on Rugosity and response distances
```{r}
R.AD <- lm(data = All.field.num, log(P_AD.m.) ~ R)
summary(R.AD)

R.FID <- lm(data = All.field.num, log(P_FID.m.) ~ R)
summary(R.FID)
#not significant
```

Regression on fish density and response distances
```{r}
#add a column for fish density per bommie (#/m^2)
All.field.num$dens <- (All.field.num$Total_fish)/(All.field.num$area)

Den.AD <- lm(data = All.field.num, log(P_AD.m.)~log(dens))
summary(Den.AD)
#Multiple R-squared:  0.09889,	Adjusted R-squared:  0.08122 
#F-statistic: 5.597 on 1 and 51 DF,  p-value: 0.02183


Den.FID <- lm(data = All.field.num, log(P_FID.m.)~log(dens))
summary(Den.FID)
#Multiple R-squared:  0.2483,	Adjusted R-squared:  0.2336 
#F-statistic: 16.85 on 1 and 51 DF,  p-value: 0.0001462
```

Plots
```{r}
#AD ~ Dens

ggplot(All.field.num, aes(log(dens), log(P_AD.m.))) + geom_point(size=3) + geom_smooth(method = "lm", se = FALSE, color = "black")+ ylab(bquote(log(Alert~distance (m)))) + xlab(bquote(log(Fish~density~(individuals/m^2)))) + theme(panel.border = element_rect(colour = "black", fill=NA, linewidth = 1), panel.background = element_rect(fill = 'white')) + annotate("text", x = 5, y = -.55, label = "paste(r ^ 2, \" = 0.081\")", parse = TRUE) + annotate("text", x = 5, y = -.7, label = "paste(p, \" = 0.022\")", parse = TRUE)


#FID ~ Dens

ggplot(All.field.num, aes(log(dens), log(P_FID.m.))) + geom_point(size=3) + geom_smooth(method = "lm", se = FALSE, color = "black")+ ylab(bquote(log(Flight~initiation~distance~(m)))) + xlab(bquote(log(Fish~density~(individuals/m^2)))) + theme(panel.border = element_rect(colour = "black", fill=NA, linewidth = 1), panel.background = element_rect(fill = 'white')) + annotate("text", x = 5, y = -.3, label = "paste(r ^ 2, \" = 0.234\")", parse = TRUE) + annotate("text", x = 5.05, y = -.45, label = "paste(p, \" = 1.46e-4\")", parse = TRUE)
```



These relationships are statistically significant, next look at density and fractal dimension/rugosity

```{r}
FD.Den <- lm(data = All.field.num, log(dens)~log(D))
summary(FD.Den)
#Multiple R-squared:  0.3881,	Adjusted R-squared:  0.3761 
#F-statistic: 32.35 on 1 and 51 DF,  p-value: 6.305e-07

R.Den <- lm(data = All.field.num, log(dens)~log(R))
summary(R.Den)
#not significant
```

Plots
```{r}
#Dens ~ D

ggplot(All.field.num, aes(D, log(dens))) + geom_point(size=3) + geom_smooth(method = "lm", se = FALSE, color = "black")+ ylab(bquote(log(Fish~density~(individuals/m^2)))) + xlab(bquote(Fractal~dimension~(D))) + theme(panel.border = element_rect(colour = "black", fill=NA, linewidth = 1), panel.background = element_rect(fill = 'white')) + annotate("text", x = 2.91, y = 5.15, label = "paste(r ^ 2, \" = 0.376\")", parse = TRUE) + annotate("text", x = 2.913, y = 4.9, label = "paste(p, \" = 6.31e-7\")", parse = TRUE)

```


Power analyses
```{r}
library(pwr)

#FRACTAL DIMENSION BETWEEN N & S  
#effect size
es_D <- (2.786604-2.746227)/0.01541155

#t.test power (unequal sample sizes)
D_power_test <- pwr.t2n.test(n1 = 31, n2 = 36, d = es_D, sig.level = 0.05, power = NULL)

D_power <- D_power_test$power
D_power
#Power (1 - P(Type II error)) is 1 => good statistical power?



#RESPONSE DISTANCE BETWEEN PRED AND NON-PRED MODELS

#Alert distance
#effect size
es_AD <- (0.2601887-0.1793878)/0.01837856

#t.test power (equal sample sizes)
AD_pwr_t <- pwr.t2n.test(n1 = 53, n2 = 48, d = es_AD, sig.level = 0.05, power = NULL)

AD_pwr <- AD_pwr_t$power
AD_pwr
#Power (1 - P(Type II error)) is 1 => good statistical power?


#Flight initiation distance
#omit NA values
P_FID_noNA <- na.omit(All_field$P_FID.m.) #n = 53
NP_FID_noNA <- na.omit(All_field$NP_FID.m.) #n = 48

#Data prep
mean_FID <- All_field %>% select(NP_FID.m., P_FID.m.) %>% summarise_if(is.numeric, mean, na.rm = TRUE)
names(mean_FID)[colnames(mean_FID) == "NP_FID.m."] <- "Non-predator"
names(mean_FID)[colnames(mean_FID) == "P_FID.m."] <- "Predator"

rotated_FID_df <- pivot_longer(mean_FID, cols = everything(), names_to = "Model_type", values_to = "Mean_dist")

SE_NP_FID <- sd(All_field$NP_FID.m., na.rm = TRUE)/sqrt(sum(!is.na(All_field$NP_FID.m.)))

SE_P_FID <- sd(All_field$P_FID.m., na.rm = TRUE)/sqrt(sum(!is.na(All_field$P_FID.m.)))

SE_FID <- data.frame(SE = c(SE_NP_FID, SE_P_FID))

sum_FID <- rotated_FID_df

sum_FID$SE <- SE_FID$SE

#Power analysis:
es_FID <- (0.4283019-0.3070833)/0.02270805
FID_pwr_t <- pwr.t2n.test(n1 = 53, n2 = 48, d = es_FID, sig.level = 0.05, power = NULL)

FID_pwr <- FID_pwr_t$power
FID_pwr
##Power (1 - P(Type II error)) is 1 => good statistical power?

#REGRESSION ON FISH DENSITY AND FRACTAL DIMENSION
#Multiple R-squared:  0.3881,	Adjusted R-squared:  0.3761 
#F-statistic: 32.35 on 1 and 51 DF,  p-value: 6.305e-07
f2_DFD <- ((0.3881)/(1-0.3881))

DFD_pwr_t <- pwr.f2.test(u = 1, v = 51, f2 = f2_DFD, sig.level = 0.05, power = NULL)
DFD_pwr <- DFD_pwr_t$power
DFD_pwr
##Power (1 - P(Type II error)) is 0.999 => good statistical power?

#REGRESSION ON FID AND FRACTAL DIMENSION
#relationship between fractal dimension and Flight Initiation Distance
#Multiple R-squared:  0.1171,	Adjusted R-squared:  0.09978 
#F-statistic: 6.764 on 1 and 51 DF,  p-value: 0.01214

f2_FIDD <- ((0.1171)/(1-0.1171))

FIDD_pwr_t <- pwr.f2.test(u = 1, v = 51, f2 = f2_FIDD, sig.level = 0.05, power = NULL)
FIDD_pwr <- FIDD_pwr_t$power
FIDD_pwr
##Power (1 - P(Type II error)) is 0.74 => good statistical power?

#REGRESSON ON FID AND FISH DENSITY
#Multiple R-squared:  0.2483,	Adjusted R-squared:  0.2336 
#F-statistic: 16.85 on 1 and 51 DF,  p-value: 0.0001462

f2_FIDdens <- ((0.2483)/(1-0.2483))

FIDdens_pwr_t <- pwr.f2.test(u = 1, v = 51, f2 = f2_FIDdens, sig.level = 0.05, power = NULL)
FIDdens_pwr <- FIDdens_pwr_t$power
FIDdens_pwr
##Power (1 - P(Type II error)) is 0.98 => good statistical power?


#REGRESSON ON AD AND FISH DENSITY
#Multiple R-squared:  0.09889,	Adjusted R-squared:  0.08122 
#F-statistic: 5.597 on 1 and 51 DF,  p-value: 0.02183

f2_ADdens <- ((0.09889)/(1-0.09889))

ADdens_pwr_t <- pwr.f2.test(u = 1, v = 51, f2 = f2_ADdens, sig.level = 0.05, power = NULL)
ADdens_pwr <- ADdens_pwr_t$power
ADdens_pwr
##Power (1 - P(Type II error)) is 0.66 => medium-high statistical power?

```

